---
title: "Hands-on Ex3 : Processing Visualising Flow Data"
author: "Flora Phyo Zin Htet"
---

## **Overview**

In this hands-on exercise, learn how to build an OD matrix.

## **Getting Started**

The code chunk below install and load **sf** and **tidyverse** packages into R environment.

```{r}
pacman::p_load(tmap, sf, DT, stplanr,
               performance,
               ggpubr, tidyverse)
```

## **Preparing the Flow Data**

### **Importing the OD data**

Firstly, we will import the *Passenger Volume by Origin Destination Bus Stops* data set downloaded from LTA DataMall by using `read_csv()` of **readr** package.

```{r}
odbus <- read_csv("data/aspatial/origin_destination_bus_202310.csv")
```

Let use display the *odbus* tibble data table by using the code chunk below.

```{r}
glimpse(odbus)
```

A quick check of odbus tibble data frame shows that the values in OROGIN_PT_CODE and DESTINATON_PT_CODE are in numeric data type. Hence, the code chunk below is used to convert these data values into character data type.

```{r}
odbus$ORIGIN_PT_CODE <- as.factor(odbus$ORIGIN_PT_CODE)
odbus$DESTINATION_PT_CODE <- as.factor(odbus$DESTINATION_PT_CODE) 
```

### **Extracting the study data**

For the purpose of this exercise, we will extract commuting flows on weekday and between 6 and 9 o\'clock.

```{r}
odbus6_9 <- odbus %>%
  filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 6 &
           TIME_PER_HOUR <= 9) %>%
  group_by(ORIGIN_PT_CODE,
           DESTINATION_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))
```

```{r}
datatable(odbus6_9)
```

We will save the output in rds format for future used

```{r}
write_rds(odbus6_9, "data/rds/odbus6_9.rds")
```

The code chunk below will be used to import the save odbus6_9.rds into R environment.

```{r}
odbus6_9 <- read_rds("data/rds/odbus6_9.rds")
```

## **Working with Geospatial Data**

For the purpose of this exercise, two geospatial data will be used. They are:

-   BusStop: This data provides the location of bus stop as at last quarter of 2022.

-   MPSZ-2019: This data provides the sub-zone boundary of URA Master Plan 2019.

Both data sets are in ESRI shapefile format.

### **Importing geospatial data**

Two geospatial data will be used in this exercise, they are:

```{r}
busstop <- st_read(dsn = "data/geospatial",
                   layer = "BusStop") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz <- st_read(dsn = "data/geospatial",
                   layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz
```

## **Geospatial data wrangling**

### **Combining Busstop and mpsz**

Code chunk below populates the planning subzone code (i.e. SUBZONE_C) of mpsz sf data frame into busstop sf data frame.

```{r}
busstop_mpsz <- st_intersection(busstop, mpsz) %>%
  select(BUS_STOP_N, SUBZONE_C) %>%
  st_drop_geometry()
```

```{r}
datatable(busstop_mpsz)
```

Before moving to the next step, it is wise to save the output into rds format.

```{r}
write_rds(busstop_mpsz, "data/rds/busstop_mpsz.rds")  
```

Next, we are going to append the planning subzone code from busstop_mpsz data frame onto odbus6_9 data frame.

```{r}
od_data <- left_join(odbus6_9 , busstop_mpsz,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_SZ = SUBZONE_C,
         DESTIN_BS = DESTINATION_PT_CODE)
```

Before continue, it is a good practice for us to check for duplicating records.

```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

If duplicated records are found, the code chunk below will be used to retain the unique records.

```{r}
od_data <- unique(od_data)
```

It will be a good practice to confirm if the duplicating records issue has been addressed fully.

Next, we will update od_data data frame cwith the planning subzone codes.

```{r}
od_data <- left_join(od_data , busstop_mpsz,
            by = c("DESTIN_BS" = "BUS_STOP_N")) 
```

```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

```{r}
od_data <- unique(od_data)
```

```{r}
od_data <- od_data %>%
  rename(DESTIN_SZ = SUBZONE_C) %>%
  drop_na() %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>%
  summarise(MORNING_PEAK = sum(TRIPS))
```

It is time to save the output into an rds file format.

```{r}
write_rds(od_data, "data/rds/od_data.rds")
```

```{r}
od_data <- read_rds("data/rds/od_data.rds")
```

## **Visualising Spatial Interaction**

In this section, you will learn how to prepare a desire line by using **stplanr** package.

### **Removing intra-zonal flows**

We will not plot the intra-zonal flows. The code chunk below will be used to remove intra-zonal flows.

```{r}
od_data1 <- od_data[od_data$ORIGIN_SZ!=od_data$DESTIN_SZ,]
```

### **Creating desire lines**

In this code chunk below, `od2line()` of **stplanr** package is used to create the desire lines.

```{r}
flowLine <- od2line(flow = od_data1, 
                    zones = mpsz,
                    zone_code = "SUBZONE_C")
```

### **Visualising the desire lines**

To visualise the resulting desire lines, the code chunk below is used.

```{r}
tmap_options(check.and.fix = TRUE) 

tm_shape(mpsz) +
  tm_polygons() +
flowLine %>%  
tm_shape() +
  tm_lines(lwd = "MORNING_PEAK",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           alpha = 0.3)
```

When the flow data are very messy and highly skewed like the one shown above, it is wiser to focus on selected flows, for example flow greater than or equal to 5000 as shown below.

```{r}
tmap_options(check.and.fix = TRUE) 
tm_shape(mpsz) +
  tm_polygons() +
flowLine %>%  
  filter(MORNING_PEAK >= 5000) %>%
tm_shape() +
  tm_lines(lwd = "MORNING_PEAK",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           alpha = 0.3)
```
