---
title: "Hands-on Exercise 2 : Global Measures of Spatial Autocorrelation"
author: "Flora Phyo Zin Htet"
---

## Overview

In this hands-on exercise, we explore how to compute Global and Local Measure of Spatial Autocorrelation (GLSA) by using **spdep** package.

## **Getting Started**

The code chunk below install and load **spdep**, **sf**, **tmap** and **tidyverse** packages into R environment.

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse)
```

## **Getting the Data Into R Environment**

### **Import shapefile into r environment**

The code chunk below uses [*st_read()*](https://r-spatial.github.io/sf/reference/st_read.html) of **sf** package to import Hunan shapefile into R.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### **Import csv file into r environment**

Import *Hunan_2012.csv* into R by using *read_csv()* of **readr** package

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

### **Performing relational join**

The code chunk below will be used to update the attribute table of *hunan*'s SpatialPolygonsDataFrame with the attribute fields of *hunan2012* dataframe. 

```{r}
hunan <- left_join(hunan,hunan2012) %>%
  select(1:4, 7, 15)
```

### **Visualising Regional Development Indicator**

prepare a basemap and a choropleth map showing the distribution of GDPPC 2012 by using *qtm()* of **tmap** package.

```{r}
equal <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal interval classification")

quantile <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal quantile classification")

tmap_arrange(equal, 
             quantile, 
             asp=1, 
             ncol=2)
```

## **Global Spatial Autocorrelation**

### **Computing Contiguity Spatial Weights**

In the code chunk below, [*poly2nb()*](https://r-spatial.github.io/spdep/reference/poly2nb.html) of **spdep** package is used to compute contiguity weight matrices for the study area. This function builds a neighbours list based on regions with contiguous boundaries. If you look at the documentation you will see that you can pass a "queen" argument that takes TRUE or FALSE as options. If you do not specify this argument the default is set to TRUE, that is, if you don't specify queen = FALSE this function will return a list of first order neighbours using the Queen criteria.

More specifically, the code chunk below is used to compute Queen contiguity weight matrix.

```{r}
wm_q <- poly2nb(hunan, 
                queen=TRUE)
summary(wm_q)
```

The summary report above shows that there are 88 area units in Hunan. The most connected area unit has 11 neighbours. There are two area units with only one neighbours.

### **Row-standardised weights matrix**

Next, we need to assign weights to each neighboring polygon. In our case, each neighboring polygon will be assigned equal weight (style="W"). This is accomplished by assigning the fraction 1/(#ofneighbors) to each neighboring county then summing the weighted income values. While this is the most intuitive way to summaries the neighbors' values it has one drawback in that polygons along the edges of the study area will base their lagged values on fewer polygons thus potentially over- or under-estimating the true nature of the spatial autocorrelation in the data. For this example, we'll stick with the style="W" option for simplicity's sake but note that other more robust options are available, notably style="B".

```{r}
rswm_q <- nb2listw(wm_q, 
                   style="W", 
                   zero.policy = TRUE)
rswm_q
```

### **Global Spatial Autocorrelation: Moran's I**

### **Maron's I test**

The code chunk below performs Moran's I statistical testing using [*moran.test()*](https://r-spatial.github.io/spdep/reference/moran.test.html) of **spdep**.

```{r}
moran.test(hunan$GDPPC, 
           listw=rswm_q, 
           zero.policy = TRUE, 
           na.action=na.omit)
```

Statistically significant, P-value 1.095e-06 is less than significance level of 5%.

#### Computing Monte Carlo Moran's I

The code chunk below performs permutation test for Moran's I statistic by using [*moran.mc()*](https://r-spatial.github.io/spdep/reference/moran.mc.html) of **spdep**. A total of 1000 simulation will be performed.

```{r}
set.seed(1234)
bperm= moran.mc(hunan$GDPPC, 
                listw=rswm_q, 
                nsim=999, 
                zero.policy = TRUE, 
                na.action=na.omit)
bperm
```

Statistically significant, P-value 0.001 is less than significance level of 5%.

#### Visualising Monte Carlo Moran's I

In the code chunk below [*hist()*](https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/hist) and [*abline()*](https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/abline) of R Graphics are used.

```{r}
mean(bperm$res[1:999])
```

```{r}
var(bperm$res[1:999])
```

```{r}
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, 
     freq=TRUE, 
     breaks=20, 
     xlab="Simulated Moran's I")
abline(v=0, 
       col="red") 
```

Use ggplot2 package

```{r}
library(ggplot2)

ggplot(data.frame(Moran_I = bperm$res), aes(x = Moran_I)) +
  geom_histogram(fill = "grey", color = "black") +
  labs(x = "Simulated Moran's I") +
  geom_vline(xintercept = 0, color = "red", linetype = "solid")
```

### **Global Spatial Autocorrelation: Geary's**

#### Geary's C test

The code chunk below performs Geary's C test for spatial autocorrelation by using [*geary.test()*](https://r-spatial.github.io/spdep/reference/geary.test.html) of **spdep**.

```{r}
geary.test(hunan$GDPPC, listw=rswm_q)
```

Statistically significant, P-value 0.0001526 is less than significance level of 5%.

#### Computing Monte Carlo Geary's C

The code chunk below performs permutation test for Geary's C statistic by using [*geary.mc()*](https://r-spatial.github.io/spdep/reference/geary.mc.html) of **spdep**.

```{r}
set.seed(1234)
bperm=geary.mc(hunan$GDPPC, 
               listw=rswm_q, 
               nsim=999)
bperm
```

Statistically significant, P-value 0.0001 is less than significance level of 5%.

#### Visualising the Monte Carlo Geary's C

Plot a histogram to reveal the distribution of the simulated values by using the code chunk below.

```{r}
mean(bperm$res[1:999])
```

```{r}
var(bperm$res[1:999])
```

```{r}
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, freq=TRUE, breaks=20, xlab="Simulated Geary c")
abline(v=1, col="red") 
```

## **Spatial Correlogram**

### **Compute Moran's I correlogram**

In the code chunk below, [*sp.correlogram()*](https://r-spatial.github.io/spdep/reference/sp.correlogram.html) of **spdep** package is used to compute a 6-lag spatial correlogram of GDPPC. The global spatial autocorrelation used in Moran's I. 

```{r}
MI_corr <- sp.correlogram(wm_q, 
                          hunan$GDPPC, 
                          order=6, 
                          method="I", 
                          style="W")
plot(MI_corr)
```

By plotting the output might not allow us to provide complete interpretation. This is because not all autocorrelation values are statistically significant. Hence, it is important for us to examine the full analysis report by printing out the analysis results as in the code chunk below.

```{r}
print(MI_corr)
```

### **Compute Geary's C correlogram and plot**

In the code chunk below, *sp.correlogram()* of **spdep** package is used to compute a 6-lag spatial correlogram of GDPPC. The global spatial autocorrelation used in Geary's C. The **plot()** of base Graph is then used to plot the output.

```{r}
GC_corr <- sp.correlogram(wm_q, 
                          hunan$GDPPC, 
                          order=6, 
                          method="C", 
                          style="W")
plot(GC_corr)
```

Similar to the previous step, we will print out the analysis report by using the code chunk below.

```{r}
print(GC_corr)
```
