---
title: "Hands-on Exercise 2 : Spatial Weights and Applications"
author: "Flora Phyo Zin Htet"
---

## Overview

In this hands-on exercise, we explore how to compute spatial weights using R.

## **Getting Started**

The code chunk below install and load **spdep**, **sf**, **tmap** and **tidyverse** packages into R environment

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, knitr)
```

###  **Import shapefile into r environment**

The code chunk below uses [*st_read()*](https://r-spatial.github.io/sf/reference/st_read.html) of **sf** package to import Hunan shapefile into R

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### **Import csv file into r environment**

The code chunk below will import *Hunan_2012.csv* into R by using *read_csv()* of **readr** package.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

### **Performing relational join**

The code chunk below will import update the attribute table of *hunan*\'s SpatialPolygonsDataFrame with the attribute fields of *hunan2012* dataframe.

```{r}
hunan <- left_join(hunan,hunan2012)%>%
  select(1:4, 7, 15)
```

## **Visualising Regional Development Indicator**

Prepare a basemap and a choropleth map showing the distribution of GDPPC 2012.

```{r}
basemap <- tm_shape(hunan) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```

## **Computing Contiguity Spatial Weights**

The code chunk below is used to compute Queen contiguity weight matrix.

```{r}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```

The summary report above shows that there are 88 area units in Hunan. The most connected area unit has 11 neighbours. There are two area units with only one heighbours.

For each polygon in our polygon object, *wm_q* lists all neighboring polygons

```{r}
wm_q[[1]]
```

Polygon 1 has 5 neighbors. The numbers represent the polygon IDs as stored in hunan SpatialPolygonsDataFrame class.

Retrive the county name of Polygon ID=1 by using the code chunk below:

```{r}
hunan$County[1]
```

The output reveals that Polygon ID=1 is Anxiang county.

Reveal the county names of the five neighboring polygons, the code chunk will be used:

```{r}
hunan$NAME_3[c(2,3,4,57,85)]
```

Retrieve the GDPPC of these five countries by using the code chunk below.

```{r}
nb1 <- wm_q[[1]]
nb1 <- hunan$GDPPC[nb1]
nb1
```

Printed output above shows that the GDPPC of the five nearest neighbours based on Queen\'s method are 20981, 34592, 24473, 21311 and 22879 respectively.

```{r}
str(wm_q)
```

### **Creating (ROOK) contiguity based neighbours**

The code chunk below is used to compute Rook contiguity weight matrix.

```{r}
wm_r <- poly2nb(hunan, queen=FALSE)
summary(wm_r)
```

The summary report above shows that there are 88 area units in Hunan. The most connect area unit has 10 neighbours. There are two area units with only one heighbours.

### **Visualising contiguity weights**

Get our longitude values we map the st_centroid function over the geometry column of us.bound and access the longitude value through double bracket notation \[\[\]\] and 1. This allows us to get only the longitude, which is the first value in each centroid.

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
```

Do the same for latitude with one key difference. We access the second value per each centroid with \[\[2\]\].

```{r}
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
```

Now that we have latitude and longitude, we use cbind to put longitude and latitude into the same object.

```{r}
coords <- cbind(longitude, latitude)
```

Check the first few observations to see if things are formatted correctly.

```{r}
head(coords)
```

#### Plotting Queen contiguity based neighbours map

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
```

####  Plotting Rook contiguity based neighbours map

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

#### Plotting both Queen and Rook contiguity based neighbours maps

```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey", main="Queen Contiguity")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
plot(hunan$geometry, border="lightgrey", main="Rook Contiguity")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## **Computing distance based neighbours**

### **Determine the cut-off distance**

determine the upper limit for distance band by using the code below:

```{r}
#coords <- coordinates(hunan)
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

The summary report shows that the largest first nearest neighbour distance is 61.79 km, so using this as the upper threshold gives certainty that all units will have at least one neighbour.

### **Computing fixed distance weight matrix**

Compute the distance weight matrix by using *dnearneigh()* as shown in the code chunk below.

```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

"Average number of links : 3.681818" is the average connectivity (number of links) for each observation in specified neighborhood distance of 62 units.

Use *str()* to display the content of wm_d62 weight matrix.

```{r}
str(wm_d62)
```

Another way to display the structure of the weight matrix is to combine [*table()*](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/table) and [*card()*](https://r-spatial.github.io/spdep/reference/card.html) of spdep.

```{r}
table(hunan$County, card(wm_d62))
```

```{r}
n_comp <- n.comp.nb(wm_d62)
n_comp$nc
```

```{r}
table(n_comp$comp.id)
```

#### Plotting fixed distance weight matrix

Plot the distance weight matrix by using the code chunk below

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE)
plot(k1, coords, add=TRUE, col="red", length=0.08)
```

The red lines show the links of 1st nearest neighbours and the black lines show the links of neighbours within the cut-off distance of 62km.

Alternatively, we can plot both of them next to each other by using the code chunk below.

```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey", main="1st nearest neighbours")
plot(k1, coords, add=TRUE, col="red", length=0.08)
plot(hunan$geometry, border="lightgrey", main="Distance link")
plot(wm_d62, coords, add=TRUE, pch = 19, cex = 0.6)
```

### **Computing adaptive distance weight matrix**

Control the numbers of neighbours directly using k-nearest neighbours, either accepting asymmetric neighbours or imposing symmetry as shown in the code chunk below.

```{r}
knn6 <- knn2nb(knearneigh(coords, k=6))
knn6
```

Similarly, we can display the content of the matrix by using *str()*.

```{r}
str(knn6)
```

Notice that each county has six neighbours, no less no more!

#### Plotting distance based neighbours

Plot the weight matrix using the code chunk below.

```{r}
plot(hunan$geometry, border="lightgrey")
plot(knn6, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## **Weights based on IDW**

Compute the distances between areas by using [*nbdists()*](https://r-spatial.github.io/spdep/reference/nbdists.html) of **spdep**.

```{r}
dist <- nbdists(wm_q, coords, longlat = TRUE)
ids <- lapply(dist, function(x) 1/(x))
ids
```

### **Row-standardised weights matrix**

Next, we need to assign weights to each neighboring polygon. In our case, each neighboring polygon will be assigned equal weight (style=\"W\"). This is accomplished by assigning the fraction 1/(#ofneighbors) to each neighboring county then summing the weighted income values. While this is the most intuitive way to summaries the neighbors\' values it has one drawback in that polygons along the edges of the study area will base their lagged values on fewer polygons thus potentially over- or under-estimating the true nature of the spatial autocorrelation in the data. For this example, we\'ll stick with the style=\"W\" option for simplicity\'s sake but note that other more robust options are available, notably style=\"B\".

```{r}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q
```

The zero.policy=TRUE option allows for lists of non-neighbors. This should be used with caution since the user may not be aware of missing neighbors in their dataset however, a zero.policy of FALSE would return an error.

See the weight of the first polygon\'s eight neighbors type:

```{r}
rswm_q$weights[10]
```

Each neighbor is assigned a 0.125 of the total weight. This means that when R computes the average neighboring income values, each neighbor\'s income will be multiplied by 0.2 before being tallied.

Using the same method, we can also derive a row standardised distance weight matrix by using the code chunk below.

```{r}
rswm_ids <- nb2listw(wm_q, glist=ids, style="B", zero.policy=TRUE)
rswm_ids
```

```{r}
rswm_ids$weights[1]
```

```{r}
summary(unlist(rswm_ids$weights))
```

## **Application of Spatial Weight Matrix**

### **Spatial lag with row-standardized weights**

Compute the average neighbor GDPPC value for each polygon. These values are often referred to as **spatially lagged values**.

```{r}
GDPPC.lag <- lag.listw(rswm_q, hunan$GDPPC)
GDPPC.lag
```

Recalled in the previous section, we retrieved the GDPPC of these five countries by using the code chunk below.

```{r}
nb1 <- wm_q[[1]]
nb1 <- hunan$GDPPC[nb1]
nb1
```

Append the spatially lag GDPPC values onto hunan sf data frame by using the code chunk below.

```{r}
lag.list <- list(hunan$NAME_3, lag.listw(rswm_q, hunan$GDPPC))
lag.res <- as.data.frame(lag.list)
colnames(lag.res) <- c("NAME_3", "lag GDPPC")
hunan <- left_join(hunan,lag.res)
```

The following table shows the average neighboring income values (stored in the Inc.lag object) for each county.

```{r}
head(hunan)
```

Next, we will plot both the GDPPC and spatial lag GDPPC for comparison using the code chunk below.

```{r}
gdppc <- qtm(hunan, "GDPPC")
lag_gdppc <- qtm(hunan, "lag GDPPC")
tmap_arrange(gdppc, lag_gdppc, asp=1, ncol=2)
```

### **Spatial lag as a sum of neighboring values**

We can calculate spatial lag as a sum of neighboring values by assigning binary weights. This requires us to go back to our neighbors list, then apply a function that will assign binary weights, then we use glist = in the nb2listw function to explicitly assign these weights.

We start by applying a function that will assign a value of 1 per each neighbor. This is done with lapply, which we have been using to manipulate the neighbors structure throughout the past notebooks. Basically it applies a function across each value in the neighbors structure

```{r}
b_weights <- lapply(wm_q, function(x) 0*x + 1)
b_weights2 <- nb2listw(wm_q, 
                       glist = b_weights, 
                       style = "B")
b_weights2
```

With the proper weights assigned, we can use lag.listw to compute a lag variable from our weight and GDPPC.

```{r}
lag_sum <- list(hunan$NAME_3, lag.listw(b_weights2, hunan$GDPPC))
lag.res <- as.data.frame(lag_sum)
colnames(lag.res) <- c("NAME_3", "lag_sum GDPPC")
```

First, let us examine the result by using the code chunk below.

```{r}
lag_sum
```

Next, we will append the *lag_sum GDPPC* field into `hunan` sf data frame by using the code chunk below.

```{r}
hunan <- left_join(hunan, lag.res)
```

Next, we will append the *lag_sum GDPPC* field into `hunan` sf data frame by using the code chunk below.

```{r}
gdppc <- qtm(hunan, "GDPPC")
lag_sum_gdppc <- qtm(hunan, "lag_sum GDPPC")
tmap_arrange(gdppc, lag_sum_gdppc, asp=1, ncol=2)
```

### **Spatial window average**

Add the diagonal element to the neighbour list, we just need to use *include.self()* from **spdep**.

```{r}
wm_qs <- include.self(wm_q)
```

Notice that the Number of nonzero links, Percentage nonzero weights and Average number of links are 536, 6.921488 and 6.090909 respectively as compared to wm_q of 448, 5.785124 and 5.090909

Let us take a good look at the neighbour list of area \[1\] by using the code chunk below.

```{r}
wm_qs[[1]]
```

Notice that now \[1\] has six neighbours instead of five.

Now we obtain weights with *nb2listw()*

```{r}
wm_qs <- nb2listw(wm_qs)
wm_qs
```

Again, we use *nb2listw()* and *glist()* to explicitly assign weight values.

Lastly, we just need to create the lag variable from our weight structure and GDPPC variable.

```{r}
lag_w_avg_gpdpc <- lag.listw(wm_qs, 
                             hunan$GDPPC)
lag_w_avg_gpdpc
```

Convert the lag variable listw object into a data.frame by using *as.data.frame()*.

```{r}
lag.list.wm_qs <- list(hunan$NAME_3, lag.listw(wm_qs, hunan$GDPPC))
lag_wm_qs.res <- as.data.frame(lag.list.wm_qs)
colnames(lag_wm_qs.res) <- c("NAME_3", "lag_window_avg GDPPC")
```

Next, the code chunk below will be used to append *lag_window_avg GDPPC* values onto *hunan* sf data.frame by using *left_join()* of **dplyr** package.

```{r}
hunan <- left_join(hunan, lag_wm_qs.res)
```

To compare the values of lag GDPPC and Spatial window average, `kable()` of Knitr package is used to prepare a table using the code chunk below.

```{r}
hunan %>%
  select("County", 
         "lag GDPPC", 
         "lag_window_avg GDPPC") %>%
  kable()
```

Lastly, *qtm()* of **tmap** package is used to plot the lag_gdppc and w_ave_gdppc maps next to each other for quick comparison.

```{r}
w_avg_gdppc <- qtm(hunan, "lag_window_avg GDPPC")
tmap_arrange(lag_gdppc, w_avg_gdppc, asp=1, ncol=2)
```

### **Spatial window sum**

Add the diagonal element to the neighbour list, we just need to use *include.self()* from **spdep**.

```{r}
wm_qs <- include.self(wm_q)
wm_qs
```

Assign binary weights to the neighbour structure that includes the diagonal element.

```{r}
b_weights <- lapply(wm_qs, function(x) 0*x + 1)
b_weights[1]
```

Notice that now \[1\] has six neighbours instead of five.

Again, we use *nb2listw()* and *glist()* to explicitly assign weight values.

```{r}
b_weights2 <- nb2listw(wm_qs, 
                       glist = b_weights, 
                       style = "B")
b_weights2
```

With our new weight structure, we can compute the lag variable with *lag.listw()*.

```{r}
w_sum_gdppc <- list(hunan$NAME_3, lag.listw(b_weights2, hunan$GDPPC))
w_sum_gdppc
```

Next, we will convert the lag variable listw object into a data.frame by using *as.data.frame()*.

```{r}
w_sum_gdppc.res <- as.data.frame(w_sum_gdppc)
colnames(w_sum_gdppc.res) <- c("NAME_3", "w_sum GDPPC")
```

Next, the code chunk below will be used to append *w_sum GDPPC* values onto *hunan* sf data.frame by using *left_join()* of **dplyr** package.

```{r}
hunan <- left_join(hunan, w_sum_gdppc.res)
```

Compare the values of lag GDPPC and Spatial window average, `kable()` of Knitr package is used to prepare a table using the code chunk below.

```{r}
hunan %>%
  select("County", "lag_sum GDPPC", "w_sum GDPPC") %>%
  kable()
```

Lastly, *qtm()* of **tmap** package is used to plot the lag_sum GDPPC and w_sum_gdppc maps next to each other for quick comparison.

```{r}
w_sum_gdppc <- qtm(hunan, "w_sum GDPPC")
tmap_arrange(lag_sum_gdppc, w_sum_gdppc, asp=1, ncol=2)
```
